---
title: "Practical Machine Learning Project"
author: "Ribery77"
date: "5 January 2016"
output: 
  html_document: 
    fig_caption: yes
    fig_height: 6
    fig_width: 8
    keep_md: yes
---



###1.Intriduction

**Background**

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).


**Data**

The training data for this project are available here: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

**Task**

What you should submit

The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.


```{r}
rm(list = ls(all = TRUE))
# Set WD
setwd("C:\\RBook\\Practical Machine Learning\\Project")
```





###2.Clean the data and made a preprocessing.

Read the training data into a data table.

```{r}
library(data.table)

# Set the connection to internet and load the dataset.
setInternet2(TRUE)
url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
D <- fread(url)
```



Read the testing data into a data table.
```{r}
url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
DTest <- fread(url)
```


We need to find in our test dataset which variable have zero NAs.

Belt, arm, dumbbell, and forearm variables that do not have any missing values in the test dataset will be **predictor candidates**.



```{r}
isAnyMissing <- sapply(DTest, function (x) any(is.na(x) | x == ""))
isPredictor <- !isAnyMissing & grepl("belt|[^(fore)]arm|dumbbell|forearm", names(isAnyMissing))
predCandidates <- names(isAnyMissing)[isPredictor]
predCandidates
```



Bear in mind that intinial dataset we need to include only the **predictor candidates** and the outcome variable, *classe*.


```{r}
varToInclude <- c("classe", predCandidates)
D <- D[, varToInclude, with=FALSE]
dim(D)
```


```{r}
names(D)
```


Convert *classe* into factor type.
```{r}
D <- D[, classe := factor(D[, classe])]
D[, .N, classe]
```


The next step is to split data into 60% training and 40% test dataset.


```{r}
library(caret)
seed <- as.numeric(as.Date("2016-01-06"))
set.seed(seed)
#set.seed(1948)
inTrain <- createDataPartition(D$classe, p=0.6)
DTrain <- D[inTrain[[1]]]
DProbe <- D[-inTrain[[1]]]
```





Preprocess the prediction variables by centering and scaling.

```{r}
X <- DTrain[, predCandidates, with=FALSE]
preProc <- preProcess(X)
preProc
```



```{r}
XCS <- predict(preProc, X)
DTrainCS <- data.table(data.frame(classe = DTrain[, classe], XCS))
```


On the other hand, we need to apply our model to the centring and scalling to the test set.
```{r}
X <- DProbe[, predCandidates, with=FALSE]
XCS <- predict(preProc, X)
DProbeCS <- data.table(data.frame(classe = DProbe[, classe], XCS))
```



Check for near zero variance.

```{r}
nzv <- nearZeroVar(DTrainCS, saveMetrics=TRUE)
if (any(nzv$nzv)) nzv else message("No variables with near zero variance")
```



Review groups of prediction variables.
```{r}
histGroup <- function (data, regex) {
  col <- grep(regex, names(data))
  col <- c(col, which(names(data) == "classe"))
  library(reshape2)
  n <- nrow(data)
  DMelted <- melt(data[, col, with=FALSE][, rownum := seq(1, n)], id.vars=c("rownum", "classe"))
  library(ggplot2)
  ggplot(DMelted, aes(x=classe, y=value)) +
    geom_violin(aes(color=classe, fill=classe), alpha=1/2) +
#     geom_jitter(aes(color=classe, fill=classe), alpha=1/10) +
#     geom_smooth(aes(group=1), method="gam", color="black", alpha=1/2, size=2) +
    facet_wrap(~ variable, scale="free_y") +
    scale_color_brewer(palette="Spectral") +
    scale_fill_brewer(palette="Spectral") +
    labs(x="", y="") +
    theme(legend.position="none")
}
histGroup(DTrainCS, "belt")
```


```{r}
histGroup(DTrainCS, "[^(fore)]arm")
```


```{r}
histGroup(DTrainCS, "dumbbell")
```


```{r}
histGroup(DTrainCS, "forearm")
```



###3.Build a prediction model
We have to use the random forest, the out of sample error have to be small. The error will be computed using the test set sample. 

Assign the parallel clusters.

```{r}
library(parallel)
library(doParallel)

# Assign them.
cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl)
```



Set the control parameters.
```{r}
ctrl <- trainControl(classProbs=TRUE,
                     savePredictions=TRUE,
                     allowParallel=TRUE)
```



Fit model over the tuning parameters.

```{r}
method <- "rf"
system.time(trainingModel <- train(classe ~ ., data=DTrainCS, method=method))
```



Stop the clusters.
```{r}
stopCluster(cl)
```



###4.Run the model on training set


```{r}
trainingModel
```




```{r}
hat <- predict(trainingModel, DTrainCS)
confusionMatrix(hat, DTrain[, classe])
```



###5.Run the model on test set


```{r}
hat <- predict(trainingModel, DProbeCS)
confusionMatrix(hat, DProbeCS[, classe])
```



###6.Show the final model

```{r}
varImp(trainingModel)
```




```{r}
trainingModel$finalModel
```


The predicicted error rate is less than 1%.

Export training model object for later analysis.

```{r}
save(trainingModel, file="trainingModel.RData")
```



###7.Predixt on test set by Jeff Leek.

Load the training model.

```{r}
load(file="trainingModel.RData", verbose=TRUE)
```



Get predictions and evaluate the model.

```{r}
DTestCS <- predict(preProc, DTest[, predCandidates, with=FALSE])
hat <- predict(trainingModel, DTestCS)
DTest <- cbind(hat , DTest)
subset(DTest, select=names(DTest)[grep("belt|[^(fore)]arm|dumbbell|forearm", names(DTest), invert=TRUE)])
```





